---
title: "Analysis script for LDP outputs study"
author: "Jason Pither, with help from Claude (Sonnet 4.6)"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    code-fold: true
    code-tools: true
    theme: cosmo
    embed-resources: true
    df-print: paged
execute:
  cache: true
  warning: false
  message: false
---

```{r libraries}
library(here)
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(irr)
library(effectsize)
library(ggplot2)
```


# Background

This script follows the analysis plans described in the [pre-registration](https://osf.io/uyqt4/files/xjmyg) for this project.  

The main steps are: 

1. importing the rater evaluation data
2. reformatting the rater data for analyses
3. conducting an assessment of inter-rater reliability using two complementary measures: 
  - Intraclass Correlation Coefficient (ICC). 
  - Krippendorff's alpha (ordinal) 
These are reported descriptively. If ICC (average measures) < 0.60, this will be flagged as a limitation and a sensitivity analysis using per-rater scores will be added.
4. primary analysis:
  - calculate the average scores for each output
  - calculate the paired differences between LDP and matched outputs
  - conduct a permutation test on the differences
  - report also the results of a paired t-test, with Cohen's D

# Import and format rater data

Import key lookup table

```{r}
#| label: load-key
#| echo: false
#| message: false

pairs_lookup <- readr::read_csv(
  here::here("data", "data_raw", "data_private", "rater_key_final_v2.csv"),
  show_col_types = FALSE
) |>
  dplyr::select(pub_id, pair_id, group)

readr::write_csv(
  pairs_lookup,
  here::here("data", "data_processed", "pairs_lookup.csv")
)

rm(pairs_lookup)
```

Import each rater's CSV

```{r}
#| label: load-raterdata

files <- list(
  JP = here("data", "data_raw", "rater_publications_final_JP.csv"),
  DH = here("data", "data_raw", "rater_publications_final_v2_DH.csv"),
  JR = here("data", "data_raw", "rater_publications_final_v2_JR_scored.csv"),
  SE = here("data", "data_raw", "rater_publications_full-SE.csv")
)

raw <- map(files, \(f) read_csv(f, show_col_types = FALSE))
```

Check that there are 38 publications in common to all rater sheets (Note that this is 4 fewer than anticipated)

```{r}
#| label: check-num-pubs

ids_common <- reduce(map(raw, \(d) unique(d$pub_id)), intersect)

# Confirm 38 publications
stopifnot(length(ids_common) == 38)
```

Reformat to wide for rater analyses

```{r}
#| label: reformat-data

# Pull pub_id + total_score from each rater, renaming JR's "total" to match
scores <- list(
  JP = raw$JP |> dplyr::select(pub_id, JP = total_score),
  DH = raw$DH |> dplyr::select(pub_id, DH = total_score),
  JR = raw$JR |> dplyr::select(pub_id, JR = total),
  SE = raw$SE |> dplyr::select(pub_id, SE = total_score)
)

# Filter each to the 38 common pub_ids, then join into a single wide tibble
scores_wide <- purrr::map(scores, \(d) dplyr::filter(d, pub_id %in% ids_common)) |>
  purrr::reduce(dplyr::left_join, by = "pub_id")

readr::write_csv(scores_wide, here::here("data", "data_processed", "rater_scores_wide.csv"))
```

# Conduct IRR analysis

## Intra-class correlation 

```{r}
#| label: icc-analysis

rater_cols <- c("DH", "JP", "JR", "SE")

# Ratings matrix: rows = publications, columns = raters
ratings_mat <- as.matrix(scores_wide[, rater_cols])

# --- ICC: two-way mixed, absolute agreement, average of k = 4 raters ---
icc_result <- irr::icc(
  ratings_mat,
  model = "twoway",
  type  = "agreement",
  unit  = "average"
)
print(icc_result)

if (icc_result$value < 0.60) {
  warning("ICC (average measures) < 0.60 — flag as limitation; per-rater sensitivity analysis required.")
}
```

## Krippendorff's alpha


```{r}
#| label: Krip-analysis
 
# kripp.alpha() expects raters as rows, subjects as columns

kripp_result <- irr::kripp.alpha(t(ratings_mat), method = "interval")
print(kripp_result)

# --- Simple percent agreement ---
agree_result <- irr::agree(ratings_mat)
print(agree_result)
```

# Primary analysis

Calculate mean scores

```{r}
#| label: calc-mean-scores
 
# Mean FAIR score per publication (average across 4 raters)
scores_mean <- scores_wide |>
  dplyr::mutate(mean_score = rowMeans(dplyr::across(dplyr::all_of(rater_cols)))) |>
  dplyr::select(pub_id, mean_score)
```

## Get lookup table

```{r}
#| label: import-lookup

key <- readr::read_csv(
  here::here("data", "data_processed", "pairs_lookup.csv"),
  show_col_types = FALSE
)
analysis_data <- key |>
  dplyr::inner_join(scores_mean, by = "pub_id")
```

## Calculate paired differences

```{r}
#| label: prep-data

# One row per pair with paired difference 
pairs_data <- analysis_data |>
  dplyr::select(pair_id, group, mean_score) |>
  tidyr::pivot_wider(names_from = group, values_from = mean_score) |>
  dplyr::mutate(D = LDP - Comparator)

stopifnot(nrow(pairs_data) == 19)

```

## Check normality

```{r}
#| label: check-normality
# Normality check (pre-specified assumption check) 

par(mfrow = c(1, 2))
hist(pairs_data$D,
     main = "Paired differences",
     xlab = "LDP \u2212 Comparator (mean FAIR score, 0\u20134 scale)")
qqnorm(pairs_data$D); qqline(pairs_data$D, col = "red")
par(mfrow = c(1, 1))

```


## Visualize the data

```{r}
#| label: pairs-plot
#| 
analysis_data |>
  ggplot2::ggplot(ggplot2::aes(x = group, y = mean_score)) +
  ggplot2::geom_line(ggplot2::aes(group = pair_id), color = "grey") +
  ggplot2::geom_point(shape = 1, size = 1.5) +
  ggplot2::scale_x_discrete(limits = c("LDP", "Comparator")) +
  ggplot2::labs(
    x = "Group",
    y = "Mean FAIR compliance score (0\u20134)"
  ) +
  ggplot2::theme_bw()
```

## Conduct permutation test

```{r}
#| label: perm-test
 
#  Permutation test (PRIMARY)
set.seed(20260329)
n_perm   <- 1000
obs_mean <- mean(pairs_data$D)

perm_means <- replicate(n_perm, {
  signs <- sample(c(-1, 1), nrow(pairs_data), replace = TRUE)
  mean(pairs_data$D * signs)
})

p_perm <- mean(perm_means >= obs_mean)

cat("=== PRIMARY: Permutation test (one-sided, n_perm = 1000) ===\n")
cat(sprintf("  Observed mean paired difference = %.3f\n", obs_mean))
cat(sprintf("  p-value                         = %.4f\n", p_perm))
```

## Supplementary paired t-test

```{r}
#| label: paired-ttest
 
#   Paired t-test (SECONDARY) 
cat("\n=== SECONDARY: One-sided paired t-test ===\n")
t_result <- t.test(pairs_data$D, alternative = "greater", mu = 0)
print(t_result)

# --- Step 7: Effect size (Cohen's d) ---
cat("\n=== Effect size ===\n")
effectsize::cohens_d(pairs_data$D, alternative = "greater")
```

