---
title: "PRISMA-Style Publication Flow Diagram"
subtitle: "Record flow from OpenAlex retrieval to final rater set — LDP quasi-experimental study"
author: "Jason Pither"
date: today
format:
  html:
    self-contained: true
    toc: false
execute:
  echo: false
  warning: false
  message: false
---

## Purpose

This document renders a PRISMA-style flow diagram tracing LDP and comparator publications
from initial OpenAlex retrieval through all filtering and matching steps to the final
19-pair rater set.

### Data sources for all counts

| Count | Source |
|---|---|
| Filtering steps (Layers 0–3) | `LDP_find-articles/data/processed_data/private/filter_log.txt` (run 2026-03-29) |
| Matching step (Script 04) | `rater_key.csv` (21 initial pairs); `LDP_publications_filtered.csv` (22 LDP authors; 1 unmatched [name hidden]) |
| Post-hoc corrections (Scripts 06–08) | `LDP_find-articles/README.md` addendum + scripts 06–08 |

### Filtering layer descriptions

**Layer 0 — Investigator co-authorship exclusion.**
Publications co-authored by any of the five study investigators (verified by ORCID) are
removed from both pools. Only papers already in each dataset are checked — no external
investigator publication lists are fetched.

**Layer 1 — Title deduplication.**
Within-author duplicates (same Unicode-normalised title, same searched name) and
cross-author duplicates (same normalised title attributed to two different searched names)
are removed. For cross-author duplicates, the alphabetically first searched name is kept.

**Layers 2a/2b — Article type and Crossref filter.**
Retains only OpenAlex `type == "article"` records; additionally drops `is_paratext == TRUE`
and Crossref types associated with non-primary-research content (reviews, editorials,
letters, corrections, etc.). *No records were excluded by this layer in either group.*

**Layer 3 — Title keyword screen.**
Titles matching word-boundary patterns for non-primary content (`review`, `perspective`,
`commentary`, `editorial`, `meta-analysis`, `response to`, `corrigendum`, `erratum`,
`retraction`) are excluded.

---

```{r}
#| label: prisma-diagram
#| fig-cap: >
#|   PRISMA-style flow of LDP and comparator publications from OpenAlex retrieval
#|   to the final 19-pair rater set. Blue boxes = LDP track; green boxes = comparator
#|   track; yellow boxes = matching / post-hoc correction steps; red boxes = excluded
#|   or removed records; amber box = replaced (not excluded) records.

library(DiagrammeR)

# Store DOT source in a variable so the export chunk can reuse it
# without duplicating the diagram definition.
dot_src <- '
digraph prisma {

  graph [rankdir  = TB,
         splines  = ortho,
         nodesep  = 0.55,
         ranksep  = 0.55,
         fontname = "Helvetica",
         bgcolor  = "white"]

  node [fontname = "Helvetica",
        fontsize  = 9,
        style     = "filled,rounded",
        shape     = rectangle,
        margin    = "0.15,0.10",
        penwidth  = 1.2]

  edge [arrowsize = 0.7,
        penwidth  = 1.0]

  //=================================================================
  // LDP MAIN FLOW  (left column)
  //=================================================================

  ldp_r [label  = "LDP PUBLICATIONS\nRetrieved via OpenAlex\nn = 55  |  23 authors",
         fillcolor = "#D6EAF8", color = "#1A5276", width = 2.8,
         fontname  = "Helvetica-Bold"]

  ldp_0 [label  = "After investigator\nco-authorship exclusion\nn = 54",
         fillcolor = "#D6EAF8", color = "#1A5276", width = 2.8]

  ldp_1 [label  = "After title\ndeduplication\nn = 54",
         fillcolor = "#D6EAF8", color = "#1A5276", width = 2.8]

  ldp_23 [label = "After type and keyword\nscreening\nn = 53  |  22 authors eligible",
          fillcolor = "#D6EAF8", color = "#1A5276", width = 2.8]

  //=================================================================
  // COMPARATOR MAIN FLOW  (right column)
  //=================================================================

  comp_r [label  = "COMPARATOR PUBLICATIONS\nRetrieved via OpenAlex\nn = 262  |  109 authors",
          fillcolor = "#D5F5E3", color = "#1D6A39", width = 2.8,
          fontname  = "Helvetica-Bold"]

  comp_0 [label  = "After investigator\nco-authorship exclusion\nn = 258",
          fillcolor = "#D5F5E3", color = "#1D6A39", width = 2.8]

  comp_1 [label  = "After title\ndeduplication\nn = 250",
          fillcolor = "#D5F5E3", color = "#1D6A39", width = 2.8]

  comp_23 [label = "After type and keyword\nscreening\nn = 248  |  105 authors eligible",
           fillcolor = "#D5F5E3", color = "#1D6A39", width = 2.8]

  //=================================================================
  // EXCLUSION BOXES  (centre column, rows 2-4)
  //=================================================================

  e_L0 [label = "Layer 0 - Investigator co-authorship exclusion\nLDP: excluded n = 1\nComparator: excluded n = 4",
        fillcolor = "#FADBD8", color = "#922B21", width = 2.8]

  e_L1 [label = "Layer 1 - Title deduplication\nLDP: none excluded\nComparator: excluded n = 8\n  (within-author: 5;  cross-author: 3)",
        fillcolor = "#FADBD8", color = "#922B21", width = 2.8]

  e_L23 [label = "Layers 2a/2b + 3 - Type and keyword screening\nLDP: excluded n = 1  (keyword match)\nComparator: excluded n = 2  (keyword match)\n[Layers 2a/2b: none excluded in either group]",
         fillcolor = "#FADBD8", color = "#922B21", width = 2.8]

  //=================================================================
  // MATCHING STEP
  //=================================================================

  match [label = "Year- and institution-matched pairs formed  (Script 04;  seed = 20260329)\nn = 21 pairs",
         fillcolor = "#FEF9E7", color = "#7D6608", width = 6.4,
         style = "filled,rounded"]

  e_match [label = "1 LDP author unmatched\nNo comparator available at same\ninstitution and calendar year",
           fillcolor = "#FADBD8", color = "#922B21", width = 2.8]

  //=================================================================
  // POST-HOC CORRECTIONS
  //=================================================================

  drop [label = "2 complete pairs dropped - no eligible LDP replacement  (Script 06)\nPAIR006  (McGill, 2024): LDP publication was not primary research\nPAIR020  (UBC, 2024): LDP publication was a conference abstract\nn = 19 pairs",
        fillcolor = "#FEF9E7", color = "#7D6608", width = 6.4,
        style = "filled,rounded"]

  e_drop [label = "Dropped: 2 pairs\n(4 publications removed)",
          fillcolor = "#FADBD8", color = "#922B21", width = 2.8]

  replace [label = "4 comparator publications replaced within intact pairs  (Scripts 07-08)\nn = 19 pairs  (unchanged)",
           fillcolor = "#FEF9E7", color = "#7D6608", width = 6.4,
           style = "filled,rounded"]

  e_replace [label = "Script 07 - replaced (ineligible): n = 3\n  R75BCN  PAIR021: case study\n  OZ2HY7  PAIR017: corresponding-author duplicate\n  TILMW1  PAIR019: conference abstract\nScript 08 - replaced (inaccessible): n = 1\n  N3V3VY  PAIR017: journal unavailable via UBC",
             fillcolor = "#FFF3CD", color = "#7D6608", width = 2.8]

  //=================================================================
  // FINAL BOX
  //=================================================================

  final [label = "FINAL RATER SET\n19 matched pairs\n(19 LDP publications  +  19 Comparator publications)",
         fillcolor = "#82E0AA", color = "#1D6A39", width = 6.4,
         style = "filled,bold,rounded", fontsize = 11]

  //=================================================================
  // MAIN FLOW EDGES
  //=================================================================

  ldp_r  -> ldp_0
  ldp_0  -> ldp_1
  ldp_1  -> ldp_23
  ldp_23 -> match

  comp_r  -> comp_0
  comp_0  -> comp_1
  comp_1  -> comp_23
  comp_23 -> match

  match   -> drop
  drop    -> replace
  replace -> final

  //=================================================================
  // EXCLUSION EDGES  (constraint = false: no rank effect)
  // Arrows originate from the step input and point to the exclusion
  // box, which is placed at the same rank as the step output.
  //=================================================================

  ldp_r  -> e_L0   [constraint = false, style = dashed, color = "#A93226"]
  comp_r -> e_L0   [constraint = false, style = dashed, color = "#A93226"]

  ldp_0  -> e_L1   [constraint = false, style = dashed, color = "#A93226"]
  comp_0 -> e_L1   [constraint = false, style = dashed, color = "#A93226"]

  ldp_1  -> e_L23  [constraint = false, style = dashed, color = "#A93226"]
  comp_1 -> e_L23  [constraint = false, style = dashed, color = "#A93226"]

  ldp_23 -> e_match   [constraint = false, style = dashed, color = "#A93226"]
  match  -> e_drop    [constraint = false, style = dashed, color = "#A93226"]
  drop   -> e_replace [constraint = false, style = dashed, color = "#7D6608"]

  //=================================================================
  // RANK GROUPS  (force same horizontal level)
  //=================================================================

  { rank = same;  ldp_r;    comp_r                   }
  { rank = same;  ldp_0;    e_L0;    comp_0           }
  { rank = same;  ldp_1;    e_L1;    comp_1           }
  { rank = same;  ldp_23;   e_L23;   comp_23          }
  { rank = same;  match;    e_match                   }
  { rank = same;  drop;     e_drop                    }
  { rank = same;  replace;  e_replace                 }

  //=================================================================
  // INVISIBLE ORDERING EDGES  (enforce left-to-right within each rank)
  //=================================================================

  ldp_r  -> comp_r  [style = invis]

  ldp_0  -> e_L0  -> comp_0  [style = invis]
  ldp_1  -> e_L1  -> comp_1  [style = invis]
  ldp_23 -> e_L23 -> comp_23 [style = invis]

  match   -> e_match   [style = invis]
  drop    -> e_drop    [style = invis]
  replace -> e_replace [style = invis]

}
'

grViz(dot_src, height = 950, width = "100%")
```

---

```{r}
#| label: export-svg-pdf

library(DiagrammeRsvg)
library(rsvg)
library(here)

# Render to SVG string using dot_src defined in the prisma-diagram chunk above
svg_str  <- DiagrammeRsvg::export_svg(grViz(dot_src))
svg_path <- here::here("scripts", "prisma_flow.svg")
pdf_path <- here::here("scripts", "prisma_flow.pdf")

# Write SVG
writeLines(svg_str, svg_path)

# Convert the saved SVG file to PDF
rsvg::rsvg_pdf(svg = svg_path, file = pdf_path)

message("Exported:\n  ", svg_path, "\n  ", pdf_path)
```
